---
title: "Introduction"
author: "Alex Pico"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"Introduction"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette will show you how to connect to Cytoscape from R using CyREST.  It will cover the installation of required packages and demonstrate some basic functionality of CyREST, commands and r2cytoscape.

It is recommended that your run this vignette *prior* to Cytoscape Automation workshops to mitigate troubleshooting time.

### Required packages
We will use pacman to install the required packages.
```{r}
if(!("pacman" %in% installed.packages())) install.packages("pacman")
library(pacman)
p_load(RJSONIO,igraph,httr,XML)
```

### Required software
The whole point of r2cytoscape is to connect with Cytoscape. You will need to install and launch Cytoscape: 

* [Download Cytoscape](http://www.cytoscape.org/download.php) 
* Complete installation wizard
* Launch Cytoscape 

### First Contact
This function is a convenient way to verify a connection to Cytoscape. It is also helpful for logging the versions of CyREST and Cytoscape in your scripts.

```{r}
checkCytoscapeVersion()
```

### My First Network
Let's create a Cytoscape network from some basic R objects
```{r}
mynodes <- data.frame(id=c("node 0","node 1","node 2","node 3"), 
                      group=c("A","A","B","B"), # optional
                      stringsAsFactors=FALSE)
myedges <- data.frame(source=c("node 0","node 0","node 0","node 2"), 
                      target=c("node 1","node 2","node 3","node 3"),
                      interaction=c("inhibits","interacts","activates","interacts"),  # optional
                      weight=c(5,3,5,9), # optional
                      stringsAsFactors=FALSE)
network.name = "myNetwork"
collection.name = "myCollection"

network.suid <- createNetwork(mynodes,myedges,network.name,collection.name)
```

### My First Style
create style with node attribute-fill mappings and some defaults
```{r}
style.name = "myStyle"
defaults <- list(NODE_SHAPE="diamond",
                 NODE_SIZE=30,
                 EDGE_TRANSPARENCY=120,
                 NODE_LABEL_POSITION="W,E,c,0.00,0.00")
nodeLabels <- mapVisualProperty('node label','id','p')
nodeFills <- mapVisualProperty('node fill color','group','d',c("A","B"), c("#FF9900","#66AAAA"))
arrowShapes <- mapVisualProperty('Edge Target Arrow Shape','interaction','d',c("activates","inhibits","interacts"),c("Arrow","T","None"))
edgeWidth <- mapVisualProperty('edge width','weight','p')

createStyle(style.name, defaults, list(nodeLabels,nodeFills,arrowShapes,edgeWidth))
applyStyle(style.name)
```

### Switch Styles
Check out the marquee style!
```{r}
applyStyle('Marquee')
```

### Browse Available Commands and Arguments
r2cytoscape helper functions
```{r}
help(package=r2cytoscape)
```

Open swagger docs for live instances of CyREST API and CyREST-supported commands:
```{r}
openCySwagger()  # CyREST API
openCySwagger("commands")  # CyREST Commands API
```

List available commands and arguments in R. Use "help" to list top level:
```{r}
commandHelp("help")  
```

List **network** commands. Note that "help" is optional:
```{r}
commandHelp("help network")  
```

List arguments for the **network select** command:
```{r}
commandHelp("help network select") 
```

Those are the basics. You are now ready to work with some real data! See [other basic and advanced R tutorials](https://github.com/cytoscape/cytoscape-automation/tree/master/for-scripters/R).
